Towards Verified Time Balancing of Security Protocols

### 2019-01-20 Src/talk.tex
more minor changes
some more changes to talk

\title{Verified Time Balancing of Security Protocols}
\author{Donovan Crichton}
\date{January 2019}

@@

\begin{frame}[fragile]
  \frametitle{Motivation}
  \begin{itemize}
    \item ASD manually verifies vendor code with containing 
          cryptographic processes.
    \item Formal Methods: A mathematically based approach to the
            specification and verification of software.
    \item Can we reduce some of the resources ASD spends on 
            manual verification by replacing with automatic 
            verification?
    \item A case study, also work towards more secure protocols.
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{The Simplified Protocol}
    \begin{itemize}
      \item Commit messages contain hashes of 256 bit random nonce.
      \item Diffie-Hellman key exchange contains modulo arithmetic.
      \item How can we formally guarantee the timing of operations?
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Approach}
  \begin{itemize}
    \item We can use the notion of propositions as types.
    \item Dependently typed functional programming languages
            also act as theorem provers for a higher-order
                  constructive logic.
    \item We can model this protocol in such a language, Idris.
    \item We can express proofs about properties of the protocol.
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Quick Background 2 - Propositions as Types}
     \begin{table}[h!]
    \begin{tabular}{c|c|c|c}
    \textbf{Logic Term} & \textbf{Logic Symbol} & 
    \textbf{Idris Symbol} & \textbf{Idris Type} \\
    \hline
      Implication & p $\Rightarrow$ q & p \mintinline{haskell}{->} q
      & Function \\
      Conjunction & p $\land$ q & \mintinline{haskell}{(p, q)} 
      & Pair / Tuple \\
      Disjunction & p $\lor$ q & \mintinline{haskell}{Either p q}
      & Tagged Union\\
      Negation & $\lnot$ p & \mintinline{haskell}{p -> Void} &
      Void Type \\
      IFF/Eq & p $\equiv$ q, p $\iff$ q & 
        \mintinline{haskell}{(p -> q, q -> p)} 
      & Pair Arrows \\
      Universal & $\forall$ x. P x & 
      \mintinline{haskell}{p -> Type} & $\Pi$ Type \\
      Existential & $\exists$ x. P x 
      & \mintinline{haskell}{(x ** P x)} & $\Sigma$ Type \\
      \hline
       & & \mintinline{haskell}{p = q} & Type Equality
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Quick Background 3 - Idris Syntax and Values as Types}
  Building a vector type in Idris: 
  \begin{minted}{haskell}
    data Vec : Nat -> Type -> Type where
      Nil  : Vec 0 a
      (::) : (x : a) -> Vec n a -> Vec (n + 1) a
  \end{minted}
  We can parameterise types over values to capture invariants in 
  the model.
  \begin{minted}{haskell}
    append : Vec n a -> Vec m a -> Vec (n + m) a
    append Nil ys = ys
    append (x :: xs) ys = x :: append xs ys 
  \end{minted}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Building a type of Prg n}
  \begin{table}[h!]
  \begin{tabular}{c|c|c|c}
  Statement & Continuation & Result & Description \\
  \hline 
  Halt & Prg 1 & Prg 1 & Terminate \\
  AssC & Prg k & Prg (k + 1) & Asn constant. \\
  AssV & Prg k & Prg (k + 1) & Asn variable. \\
  UnOp & Prg k & Prg (k + 1) & Asn result of unary op. \\
  BinOP & Prg k & Prg (k + 1) & Asn result of binary op. \\
  Do & Prg k & Prg (m * n + k) & Run Prg m, n times. \\
  Cond & Prg k & Prg (n + k) & Branch on Prg n. \\
  \end{tabular} 
  \end{table}
  Conditionals require both branches to be Prg n. Ensuring
  That all branches of the program are correct by construction.
\end{frame}

@@

      \path [->] (Prg) edge node {} (Elab);
      \path [->] (Prg) edge node {} (Comp);
  \end{tikzpicture}
  \end{center}
  \end{minipage}
  \begin{minipage}[c]{\textwidth-6cm}
  \begin{itemize}
    \item The small, correct core language can be elaborated to
          a more full-featured language.
    \item The size of the core language makes the burden of proofs
          much lighter.
    \item The compiler can map the core language expressions down
          to something more real world (e.g C, assembler).
  \end{itemize}
  \end{minipage}
\end{frame}

>>>>

\title{Towards Verified Time Balancing of Security Protocols}
\author{Donovan Crichton}
\date{January 2019}

@@

\begin{frame}[fragile]
  \frametitle{Motivation}
  \begin{itemize}
    \item ASD manually verifies vendor code with containing 
          cryptographic processes.
    \item Formal Methods: A mathematically based approach to the
            specification and verification of software.
    \item Can we reduce some of the resources ASD spends on 
            manual verification by replacing with automatic 
            verification?
    \item Can we also further research into secure communications
          protocols?
    \item A case study.
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{The Simplified Protocol}
    \begin{itemize}
      \item Commit messages that contain SHA-256 hashes
      \item Diffie-Hellman key exchange contains modulo arithmetic.
      \item How can we formally guarantee the timing of operations?
      \item Lets cover some background before diving in!
    \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Approach}
  \begin{itemize}
    \item We can use the notion of propositions as types.
    \item Functional programming languages that support dependent
          types can act as theorem provers under a higher order
          constructive logic.
    \item We can model this protocol in such a language, Idris.
    \item We can express proofs about properties of the protocol.
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Quick Background 2 - Idris Syntax and Values as Types}
  Building a vector type in Idris: 
  \begin{minted}{haskell}
    data Vec : Nat -> Type -> Type where
      Nil  : Vec 0 a
      (::) : (x : a) -> Vec n a -> Vec (n + 1) a
  \end{minted}
  We can parameterise types over values to capture invariants in 
  the model.
  \begin{minted}{haskell}
    append : Vec n a -> Vec m a -> Vec (n + m) a
    append Nil ys = ys
    append (x :: xs) ys = x :: append xs ys 
  \end{minted}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Quick Background 3 - Propositions as Types}
  Under the assumptions of referential transparency and totality.
     \begin{table}[h!]
    \begin{tabular}{c|c|c|c}
    \textbf{Logic Term} & \textbf{Logic Symbol} & 
    \textbf{Idris Symbol} & \textbf{Idris Type} \\
    \hline
      Implication & p $\Rightarrow$ q & p \mintinline{haskell}{->} q
      & Function \\
      Conjunction & p $\land$ q & \mintinline{haskell}{(p, q)} 
      & Pair / Tuple \\
      Disjunction & p $\lor$ q & \mintinline{haskell}{Either p q}
      & Tagged Union\\
      Negation & $\lnot$ p & \mintinline{haskell}{p -> Void} &
      Void Type \\
      IFF/Eq & p $\equiv$ q, p $\iff$ q & 
        \mintinline{haskell}{(p -> q, q -> p)} 
      & Pair Arrows \\
      Universal & $\forall$ x. P x & 
      \mintinline{haskell}{p -> Type} & $\Pi$ Type \\
      Existential & $\exists$ x. P x 
      & \mintinline{haskell}{(x ** P x)} & $\Sigma$ Type \\
      \hline
       & & \mintinline{haskell}{p = q} & Type Equality
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Building a type of Prg n}
  \begin{table}[h!]
  \begin{tabular}{c|c|c|c}
  Statement & Continuation & Result & Description \\
  \hline 
  Halt & - & Prg 1 & Terminate \\
  AssC & Prg k & Prg (k + 1) & Asn constant. \\
  AssV & Prg k & Prg (k + 1) & Asn variable. \\
  UnOp & Prg k & Prg (k + 1) & Asn result of unary op. \\
  BinOP & Prg k & Prg (k + 1) & Asn result of binary op. \\
  Do & Prg k & Prg (m * n + k) & Run Prg m, n times. \\
  Cond & Prg k & Prg (n + k) & Branch on Prg n. \\
  Skip & Prg k * Prg (k + 1) & Do Nothing.
  \end{tabular} 
  \end{table}
  Conditionals require both branches to be Prg n. Ensuring
  That all branches of the program are correct by construction.
\end{frame}

@@

      \path [->] (Prg) edge node {} (Elab);
      \path [->] (Prg) edge node {} (Comp);
  \end{tikzpicture}
  \end{center}
  \end{minipage}
  \begin{minipage}[c]{\textwidth-6cm}
  \begin{itemize}
    \item The small, correct core language can be elaborated to
          a more full-featured language.
    \item The size of the core language makes the burden of proofs
          much lighter.
    \item The compiler can map the core language expressions down
          to something more ``real world" (e.g C, assembler).
  \end{itemize}
  \end{minipage}
\end{frame}
